Use R ,EDA techniques to explore a wine quality dataset by Kamaljit Chahal;
============================================================================



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code

library(ggplot2)
library(grid)
library(gridExtra)
library(GGally)
library(dplyr)
library(tidyr)
library(scales)
library(memisc)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd('C:/Users/cm641e/Desktop/Nanodegree')
wines <- read.csv("wineQualityReds.csv",  row.names = 1, 
                  stringsAsFactors = FALSE)
head(wines)
wines$quality.level <- ifelse(wines$quality < 5, "low", 
                             ifelse(wines$quality < 7, "average", "high"))
wines$quality.level <- factor(wines$quality.level, 
                             levels=c("high", "average", "low"), ordered=TRUE)
attach(wines)
wines$quality <- factor(wines$quality, ordered = TRUE)
wines$quality
glimpse(wines)

```


> **Dataset summary**: There are 13 variables( 1 additional quality.level 
created )  with 1599 observations. Quality variable is discrete.rest of the 
variable are continous quality is ranging from 3 to 8 and average quality 
of wines dataset falls around 5-6 .distribution of quality appears to be normal 
with most wines average quality (5) and lower counts at low quality as (3-4) 
and high quality(8).no wines below quality of 3 and no wines above the quality
of 8.

# Univariate Plots Section





```{r echo=FALSE, Univariate_Plots_Q}
summary(wines$quality)
ggplot(data = wines, aes(x = quality))+
geom_bar()
```


> **Quality**:quality is ranging from 3 to 8 and average quality of wines 
dataset falls around 5-6 .


#defining the function will help create the univariate plots fast .


```{r echo=FALSE, Univariate_Plots_funct}
histogram <- function(variable, ...){
  ggplot(data = wines, aes_q(as.name(variable)))+
    geom_histogram(...)
}

print_summary <- function(variable){
  print(summary(wines[[variable]]))
}

plot_and_summary <- function(variable, ...){
  print_summary(variable)
  histogram(variable, ...)
}
```


#PLots Area


```{r echo=FALSE, Univariate_Plots_FA}
plot_and_summary("fixed.acidity", binwidth = 0.2)+
  scale_x_continuous(breaks = 4:16)
```


> **Fixed Acidity**:The median fixed acidity is 7.90 . 1st quantile is  7.10 
and 3rd quantile is 9.20. The distribution of fixed acidity is slightly 
right skewed.  some outliers are above 15 .mean value is 8.32 and max 
value is 15.90.


```{r echo=FALSE, Univariate_Plots_VA}
plot_and_summary("volatile.acidity", binwidth = 0.02)+
  scale_x_continuous(breaks = seq(0, 1.6, 0.1))
```


> **volatile. acidity**:The median volatile. acidity is 0.52.Most values 
are in the range of 0.39 - 0.64 and outliers are seen around 1.0 and above.The 
distribution of volatile acidity is non-symmetric and there are two peaks at 
0.4 and 0.6. 1st quantile = 0.39 and 3rd qunatile = 0.64. max value of 1.58. 


```{r echo=FALSE, Univariate_Plots_CIA}
plot_and_summary("citric.acid", binwidth = 0.03)+
  scale_x_continuous(breaks = seq(0, 1, 0.125))
```


> **citric.acid**:citric.acid is very skewed to the right.Median is 0.260 and 
most of the readings are in the range of 0.090 - 0.420.


```{r echo=FALSE, Univariate_Plots_RS}
plot_and_summary("residual.sugar", binwidth = 0.2)+
  scale_x_continuous(breaks = 0:16)+
  coord_cartesian(xlim = c(0, 7))
```


> **residual.sugar**:residual.sugar looks normally distributed, but there 
are some outliers between 7 - 15 range.median value is 2.2. 


```{r echo=FALSE, Univariate_Plots_AL}
plot_and_summary("alcohol", binwidth = 0.2)+
  scale_x_continuous(breaks = seq(8, 15, 1))
```


> **alcohol**:alcohol is skewed to the right. most of the values in the range 
of 9.50 -11.10 . median is 10.20 .


```{r echo=FALSE, Univariate_Plots_SU}
plot_and_summary("sulphates", binwidth = 0.05)+
  scale_x_continuous(breaks = seq(0, 2, 0.25))
```


> **sulphates**:sulphates looks normally distributed, but there are some 
outliers above 1.25 .median value is 0.62 and most of the readings in the 
range of 0.55 -0.73.


```{r echo=FALSE, Univariate_Plots_PH}
plot_and_summary("pH", binwidth = 0.02)+
  scale_x_continuous(breaks = seq(0, 4.5, 0.1))
```


> **pH**:pH looks normally distributed.median value is 3.3 and most of the 
values are in the range of 3.2 - 3.4 .


```{r echo=FALSE, Univariate_Plots_DY}
plot_and_summary("density", binwidth = 0.0005)+
  scale_x_continuous(breaks = seq(0.990, 1.005, 0.0025))
```


> **density**:density looks normally distributed. median is 0.99 . most of the
observations are in the range of 0.995 - 0.997


```{r echo=FALSE, Univariate_Plots_SD}
plot_and_summary("total.sulfur.dioxide", binwidth = 5)+
  scale_x_continuous(breaks = seq(0, 300, 25))
```


> **total.sulfur.dioxide**:total.sulfur.dioxide looks skewed to the right.
median is 38. most of the readings in the range of 22- 62. there are outliers 
above 275 .


```{r echo=FALSE, Univariate_Plots_FSD}
plot_and_summary("free.sulfur.dioxide", binwidth = 2)+
  scale_x_continuous(breaks = seq(0, 70, 5))+
  scale_color_brewer(type='seq')

```


> **free.sulfur.dioxide**:free.sulfur.dioxide looks skewed to the right.
median is 14. most of the values in the range of 7 -21 . there are outliers 
around 60 .


```{r echo=FALSE, Univariate_Plots_CS}
plot_and_summary("chlorides", binwidth = 0.01)+
  scale_x_continuous(breaks = seq(0, 0.2, 0.05))+
  coord_cartesian(xlim = c(0, 0.2))+
  scale_color_brewer(type='seq')
```


> **chlorides**:chlorides look normally distributed, but there are  very 
distant outliers.median is 0.079 and most of the readings in the range of 
0.0700 - 0.09000.


# Univariate Analysis


### What is the structure of your dataset?
> **Answer**:There are 1599 red wines in the dataset with twelve variables and 
an identifier number. Except quality, which is discrete, the rest of the 
features are continuous, although most of the values of free.sulfur.dioxide 
and total.sulfur.dioxide are whole.



### What is/are the main feature(s) of interest in your dataset?
> **Answer**:The main feature is quality. 



### What other features in the dataset do you think will help support your\
investigation into your feature(s) of interest?
> **Answer**:pysciochemical characteristics of the variable determines flavor.
density represents sugar content and alochol .



### Did you create any new variables from existing variables in the dataset?
> **Answer**:yes , i created quality.level , it shows ranges from low to high .
I didnt thought of combining/ transforming any other variables that can results 
in a new one.



### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or \
change the form of the data? If so, why did you do this?
> **Answer**:There were some distributions skewed to right, no missing values 
and no need to adjust the data.some of them have outliers. 




# Bivariate Plots Section



```{r echo=FALSE, Bivariate_Plots_VA_1}
ggplot(data = wines, aes(y = volatile.acidity, x = quality)) +
  geom_boxplot()+
  scale_color_brewer(type='seq')+
  theme_dark()
```


```{r echo=FALSE, Bivariate_Plots_VA}
va1 <- ggplot(data = wines, aes(volatile.acidity, colour = quality)) +
  geom_freqpoly(binwidth = 0.05)+
  scale_color_brewer(type='seq')+
  theme_dark()

va2 <- ggplot(data = wines, aes(volatile.acidity, fill = quality)) +
  geom_histogram(binwidth = 0.05)+
  scale_color_brewer(type='seq')+
  theme_dark()

grid.arrange(va1, va2, ncol=1)
```


> **volatile acidity vs Quality**:volatile acidity increases,with quality 
decreases, two observations :
1. Wines with a quality 7 & 8 approximately same median volatile acidity. 
2. Outliers with a quality of 7 & 8 have a volatile acidity approximate similar 
median of the worst rated wines: volatile acidity can't find direct 
relation with quality.it required to be looked in multivaribate plots.


```{r echo=FALSE, Bivariate_Plots_CA}
ggplot(data = wines, aes(y = citric.acid, x = quality)) +
  geom_boxplot()+
  scale_color_brewer(type='seq')+
  theme_dark()
```

```{r echo=FALSE, Bivariate_Plots_CA_1}
ca1 <- ggplot(data = wines, aes(citric.acid, colour = quality)) +
  geom_freqpoly(binwidth = 0.05)+
  scale_color_brewer(type='seq')+
  theme_dark()

ca2 <- ggplot(data = wines, aes(citric.acid, fill = quality)) +
  geom_histogram(binwidth = 0.05)+
  scale_color_brewer(type='seq')+
  theme_dark()

grid.arrange(ca1, ca2, ncol=1)
```


> **Cirtic Acid vs Quality**:quality and citric.acid are positively correlated.
wines with a quality score of 7 or 8 have similar levels of citric acid. 
The median citric acid quantity for low quality wines is very low.


```{r echo=FALSE, Bivariate_Plots_SU}
ggplot(data = wines, aes(y = sulphates, x = quality)) +
  geom_boxplot()+
  scale_color_brewer(type='seq')+
  theme_dark()
```


```{r echo=FALSE, Bivariate_Plots_SU_1}
SU1 <- ggplot(data = wines, aes(sulphates, colour = quality)) +
  geom_freqpoly(binwidth = 0.05)+
  scale_color_brewer(type='seq')+
  theme_dark()

SU2 <- ggplot(data = wines, aes(sulphates, fill = quality)) +
  geom_histogram(binwidth = 0.05)+
  scale_color_brewer(type='seq')+
  theme_dark()

grid.arrange(SU1, SU2, ncol=1)
```


> **Sulphates vs Quality**:quality and sulphates are slightly correlated but 
not very conclusionfull .Ouliers are present.Higher sulphates concentration 
means higher quality.


```{r echo=FALSE, Bivariate_Plots_AL}
ggplot(data = wines, aes(y = alcohol, x = quality)) +
  geom_boxplot()+
scale_color_brewer(type='seq')+
  
  theme_dark()
```


```{r echo=FALSE, Bivariate_Plots_AL_1}
AL1 <- ggplot(data = wines, aes(alcohol, colour = quality)) +
  geom_freqpoly(binwidth = 0.2)+
  scale_color_brewer(type='seq')+
  theme_dark()

AL2 <- ggplot(data = wines, aes(alcohol, fill = quality)) +
  geom_histogram(binwidth = 0.2)+
  scale_color_brewer(type='seq')+
  theme_dark()

grid.arrange(AL1, AL2, ncol=1)
```


> **Alcohol vs Quality**: quality increases with alcohol content.
alcohol content outliers present around quality of 5. 
lower end of quaility  have similar alcohol content median.


```{r echo=FALSE, Bivariate_Plots_DU}
ggplot(data = wines, aes(y = density, x = quality)) +
  geom_boxplot()+
  scale_color_brewer(type='seq')+
  theme_dark()
```


```{r echo=FALSE, Bivariate_Plots_DU_1}
du1 <- ggplot(data = wines, aes(density, colour = quality)) +
  geom_freqpoly(binwidth = 0.0007)+
  scale_color_brewer(type='seq')+
  theme_dark()

du2 <- ggplot(data = wines, aes(density, fill = quality)) +
  geom_histogram(binwidth = 0.0007)+
  scale_color_brewer(type='seq')+
  theme_dark()

grid.arrange(du1, du2, ncol=1)

```


> **density vs Quality**: Lower density means higher quality. Density is
directly related to sugar and alcohol.We can check in multivibrate.It is in
fact negativately correlated.


```{r echo=FALSE, Bivariate_Plots_PH}
ggplot(data = wines, aes(y = pH, x = quality)) +
  scale_color_brewer(type='seq')+
  theme_dark()+
  geom_boxplot()
```


```{r echo=FALSE, Bivariate_Plots_PH_1}
PH1 <- ggplot(data = wines, aes(pH, colour = quality)) +
  geom_freqpoly(binwidth = 0.07)+
  scale_color_brewer(type='seq')+
  theme_dark()

PH2 <- ggplot(data = wines, aes(pH, fill = quality)) +
  geom_histogram(binwidth = 0.07)+
  scale_color_brewer(type='seq')+
  theme_dark()

grid.arrange(PH1, PH2, ncol=1)

```


> **ph vs Quality**: lower pH means higher quality.It is in
fact negativately correlated.


```{r echo=FALSE, Bivariate_Plots_TSD}
ggplot(data = wines, aes(y = total.sulfur.dioxide, x = quality)) +
  geom_boxplot()+
  scale_color_brewer(type='seq')+
  theme_dark()
```


```{r echo=FALSE, Bivariate_Plots_TSD_1}
TSD1 <- ggplot(data = wines, aes(total.sulfur.dioxide, colour = quality)) +
  geom_freqpoly(binwidth = 1.9)+
  scale_color_brewer(type='seq')+
  theme_dark()

TSD2 <- ggplot(data = wines, aes(total.sulfur.dioxide, fill = quality)) +
  geom_histogram(binwidth = 1.9)+
  scale_color_brewer(type='seq')+
  theme_dark()

grid.arrange(TSD1, TSD2, ncol=1)

```


> **Total sulfur dioxide vs Quality**: outliers around quality of 6 &7 .in the
middle higher concentration around quality of 5.



```{r echo=FALSE, Bivariate_Plots_SD}
ggplot(data = wines, aes(y = free.sulfur.dioxide, x = quality)) +
  geom_boxplot()+
  scale_color_brewer(type='seq')+
  theme_dark()
```


```{r echo=FALSE, Bivariate_Plots_SD_1}
SD1 <- ggplot(data = wines, aes(free.sulfur.dioxide, colour = quality)) +
  geom_freqpoly(binwidth = 0.9)+
  scale_color_brewer(type='seq')+
  theme_dark()

SD2 <- ggplot(data = wines, aes(free.sulfur.dioxide, fill = quality)) +
  geom_histogram(binwidth = 0.9)+
  scale_color_brewer(type='seq')+
  theme_dark()

grid.arrange(SD1, SD2, ncol=1)

```


> **Free sulfur dioxide vs Quality**: outliers around quality of 5 and above.
in the middle higher concentration around quality of 5.Lower on low end and 
high end of the quality. 


```{r echo=FALSE, Bivariate_Plots_RS}
ggplot(data = wines, aes(y = residual.sugar, x = quality)) +
  geom_boxplot()+
  scale_color_brewer(type='seq')+
  theme_dark()
```






```{r echo=FALSE, Bivariate_Plots_RS_1}
Rsu1 <- ggplot(data = wines, aes(residual.sugar , colour = quality)) +
  geom_freqpoly(binwidth = 0.3)+
  theme_dark()+
  scale_color_brewer(type='seq')

Rsu2 <- ggplot(data = wines, aes(residual.sugar , fill = quality)) +
  geom_histogram(binwidth = 0.3)+
  theme_dark()+
  scale_color_brewer(type='seq')

grid.arrange(Rsu1, Rsu2, ncol=1)

```






```{r echo=FALSE, Bivariate_Plots_RS_2}

Rsu3 <- ggplot(data = wines, aes(residual.sugar , colour = quality)) +
  coord_cartesian(xlim=c(0.6,18))+
  geom_histogram(binwidth = 0.5)+
  scale_x_continuous(breaks = seq(0,18,2))+
  theme_dark()+
  labs(x= 'Residual sugar( g / dm^3)',
       y= 'Count')

Rsu4 <- ggplot(data = wines, aes(residual.sugar , colour = quality)) +
  coord_cartesian(xlim=c(0.6,18))+
  geom_histogram(binwidth = 0.05)+
  scale_x_log10(breaks = seq(0,18,2))+
  labs(x= 'Residual sugar( g / dm^3)',
       y= 'Count')


grid.arrange(Rsu3, Rsu4, ncol=1)


```




> **Residual.sugar vs Quality**: outliers around quality of 5 and above.
Looks like residual sugar have low impact on quality . 
Looks like residual decreases/ stayed same with increasing quality in some 
cases.
Second Plot used log 10 scale for better viusualization .

```{r echo=FALSE, Bivariate_Plots_CL}
ggplot(data = wines, aes(y = chlorides, x = quality)) +
  geom_boxplot()+
  scale_color_brewer(type='seq')+
  theme_dark()
```


```{r echo=FALSE, Bivariate_Plots_CL_1}
CL1 <- ggplot(data = wines, aes(chlorides, colour = quality)) +
  geom_freqpoly(binwidth = 0.006)+
   scale_color_brewer(type='seq')+
  theme_dark()

CL2 <- ggplot(data = wines, aes(chlorides, fill = quality)) +
  scale_color_brewer(type='seq')+
  theme_dark()+
  geom_histogram(binwidth = 0.05)+
  scale_x_log10(breaks = seq(0,0.4,0.1))

grid.arrange(CL1, CL2, ncol=1)

```


> **Chlorides vs Quality**: outliers around quality of 5 and above.
Looks like Less chlorides means higher quality but it is very clearly visible.


#Finding Correlations :

```{r echo=FALSE, Bivariate_Plots_COR}
cor.test(~ fixed.acidity + volatile.acidity, data = wines)
```


> **Correlation :Fixed.acidity & Volatile.acidity**:weak negative correlation.


```{r echo=FALSE, Bivariate_Plots_COR1}
cor.test(~ residual.sugar + alcohol, data = wines)
```


> **Correlation :residual.sugar & alcohol**:there is no strong correlation 
between the alcohol content and the residual sugar, since the alcohol
comes from the fermentation process of the sugars.



```{r echo=FALSE, Bivariate_Plots_COR2}
cor.test(~ total.sulfur.dioxide + sulphates, wines)
```


> **Correlation :total.sulfur.dioxide & sulphates**:there is no strong 
correlation between the total.sulfur.dioxide & sulphates.


```{r echo=FALSE, Bivariate_Plots_COR3}
cor.test(~ free.sulfur.dioxide + sulphates, wines)
```


> **Correlation :free.sulfur.dioxide & sulphates**:there is no strong 
correlation between the free.sulfur.dioxide & sulphates.


```{r echo=FALSE, Bivariate_Plots_COR4}
cor.test(~ pH + citric.acid, data = wines)

```


> **Correlation :pH & citric.acid**:negative correlation between the pH &
citric.acid.low pH values indicates acidity.


```{r echo=FALSE, Bivariate_Plots_COR5}
cor.test(~ pH + volatile.acidity, data = wines)
```


> **Correlation :pH  & volatile.acidity**:weak correlation between the pH &
volatile.acidity.


```{r echo=FALSE, Bivariate_Plots_COR6}
cor.test(~ density + alcohol, data = wines)
```


> **Correlation :density & alcohol**: negative correlation.High content of 
alcohol are associated with low density ,since alcohol is less dense than water.



# Bivariate Analysis



### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
> **Answer**:citric.acid ,alcohol, volatile.acidity ,sulphates variable are 
strongly correlated to the quality of a wine.weak correlation with pH,density .



### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)? 
> **Answer**:relationship between Citric.acid & volatile.acidity negatively 
correlation. negative correlation between alcohol & density .High content of 
alcohol are associated with low density .



### What was the strongest relationship you found?
> **Answer**:strongest relationship is between quality & alcohol content.



# Multivariate Plots Section



```{r echo=FALSE, Multivariate_Plots}

plot1 <- ggplot(data = wines, aes(y = volatile.acidity, x = alcohol,
                                  colour=quality)) +
  scale_color_brewer(type='seq')+
  theme_dark()+
  geom_point(alpha = 0.8, size = 1)+
  geom_point()

plot2 <- ggplot(data = wines, aes(y = sulphates, x = alcohol, 
                                  colour=quality)) +
  scale_color_brewer(type='seq')+
  geom_point(alpha = 0.8, size = 1) +
  theme_dark()+
  geom_point()

plot3 <- ggplot(data = wines, aes(y = volatile.acidity, x = sulphates,
                                  colour=quality)) +
  scale_color_brewer(type='seq')+
  geom_point(alpha = 0.8, size = 1)+
  theme_dark()+
  geom_point()




grid.arrange(plot1, plot2, plot3,ncol=1)

```

> **alochol vs Voilate.acidity  **: lower Volatile acidity & higher content of 
alochol seems better wines .


> **alochol vs sulphates plot  **: Wines with higher alcohol & Sulphates content 
produce better wine.


> **sulphates plot vs volatile acidity plot  **: lower sulphates and Volatile 
acidity seems better wines.





```{r echo=FALSE, Multivariate_Plots_2}

plot4 <- ggplot(data = wines, aes(y = volatile.acidity, x = citric.acid, 
                                  colour=quality)) +
   scale_color_brewer(type='seq')+
  geom_jitter()+
  theme_dark()

plot5 <- ggplot(data = wines, aes(y = alcohol, x = citric.acid, 
                                  colour=quality)) +
   scale_color_brewer(type='seq')+
  geom_jitter()+
  theme_dark()

plot6 <- ggplot(data = wines, aes(y = sulphates, x = citric.acid, 
                                  colour=quality)) +
   scale_color_brewer(type='seq')+
  geom_jitter()+
  theme_dark()
grid.arrange(plot4,plot5,plot6,ncol=1)


```


> **Citric.acid vs Voilate.acidity  **: higher Volatile acidity & low content of 
citric acid is indicative of bad  wines .


> **alochol vs citric.acid  **: Wines with higher alcohol & higher citric.acid 
produce better wine.


> **sulphates plot vs citric.acid plot  **: higher sulphates and slightly 
higher citric acid seems better wines.




#chart.Correlation():


```{r echo=FALSE, Multivariate_Plots_3}
library("PerformanceAnalytics")
my_data <- wines[, c("citric.acid","alcohol","volatile.acidity","sulphates")]
chart.Correlation(my_data, histogram=TRUE, pch=19)

```



> **Correlation results  **:


> **Citric.acid vs Voilate.acidity  **: negative correlation 
> **alochol vs citric.acid  **: positive correlation 
> **sulphates plot vs citric.acid plot  **: positive correlation



#subsetting the data to get better insights :


```{r echo=FALSE, Multivariate_Plots_4}
mod.quality <- subset(wines, quality != 6 & quality != 5)
plot7 <- ggplot(data = mod.quality, aes(y = citric.acid, x = alcohol, 
                                        colour=quality)) +
   scale_color_brewer(type='seq')+
  geom_jitter()+
  theme_dark()

plot8 <-ggplot(data = mod.quality, aes(y = sulphates, x = alcohol, 
                                       colour=quality)) +
  scale_color_brewer(type='seq')+
  geom_jitter()+
  theme_dark()

plot9 <-ggplot(data = mod.quality, aes(y = volatile.acidity, x = alcohol, 
                                       colour=quality)) +
   scale_color_brewer(type='seq')+
  geom_jitter()+
  theme_dark()

plot10 <- ggplot(data = mod.quality, aes(y = citric.acid, x = volatile.acidity, 
                                         colour=quality)) +
   scale_color_brewer(type='seq')+
  geom_jitter()+
  theme_dark()

grid.arrange(plot7,plot8,plot9,plot10,ncol=1)

```
 
 
> **Voilate.acidity vs citric acid plot **: high volatile acidity and low 
citric acid is indicative of bad wines.


> **alochol vs sulphates plot  **: high values of alcohol and sulphates
are deciding factor  in the quality of the wines.

> **alochol vs citric acid plot **:high values of alcohol and citric.acid 
are deciding factor  in the quality of the wines.


> **alochol vs volatile acidity plot  **: bad quality wines 
tend to have more than 0.5+ of acetic acid, high quality wines are below than
0.5.



# Multivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
> **Answer **:alcohol , sulphates and volatile acidity are closely correlated 
with quality. Higher alcohol , higher sulphates  and lower acidity have higher
quality.Higher citric acid & higher alcohol gets higher quality.


### Were there any interesting or surprising interactions between features?
> **Answer **:volatile.acidity threshold at approximately 0.8-0.9 of acetic 
acid described by the authors of the dataset & can be seen in the plots also .
none of the good quality wines have more than 0.8 values on volatile.acidity.


# Final Plots and Summary



### Plot One
```{r echo=FALSE, Plot_One}
ggplot(data = wines, aes(x = quality)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 800, 50)) +
  scale_x_discrete(breaks = seq(0, 14, 1)) +
  
  ggtitle('Quality hist') +
  xlab('Quality Rating') + 
  ylab('# of wines')
```

### Description One
> **Description 1 **:wines with quality of 5 around  1300 out of  1599.no wines
with quality equal to 1 ,2 ,9,10 .


### Plot Two
```{r echo=FALSE, Plot_Two}

boxplots_dots <- function(variable){
  ggplot(data = wines, aes_q(x = ~quality, y = as.name(variable)))+
    geom_boxplot()+
    geom_jitter(width = 0.7, alpha = 1/10)+
    geom_line(aes(group = 1),
                  stat = "summary",
                  fun.y = median,
                  size = 2,
                  alpha = 1)
}

boxplots_dots("alcohol")+
  xlab("Quality Rating")+
  ylab("Alcohol (% by volume)")+
  ggtitle("alcohol vs quality")


```

### Description Two
> **Description 2 **:generally , higher alcohol level yields better quality of 
wines. quality from 5 to 8, see upwards direction trend.


### Plot Three

```{r echo=FALSE, Plot_Three}
ggplot(data = wines, aes(y = citric.acid, x = alcohol, colour=quality)) +
  scale_color_brewer(type='seq')+
  coord_cartesian(xlim=c(8,15),ylim=c(0,1))+
  geom_jitter() +
  theme_dark()+
  ggtitle('quality of wines with deciding factors') +
  xlab("Alcohol (% by volume)")+
  ylab("citric.acid- g/dm^3") 
```


### Description Three
> **Description 3 **:high quality wines have high citric acid and alcohol &  
opposite for low quality wines. 




# Reflection
> **Reflection Part**:to gain knowledge of the R plotting functionality & 
libraries to explore a winesdataset. there are not much wines on 0 to 2 , above
8 rated quality .

> ** **:domain knowledge is must to understand the data ahead of 
data analysis . This dataset is ready for machine learning as there are no 
missing data.R-bloggers correlations and libraries is a great helping hand .
This dataset should be compared with white wines to see similar trends .

> ** **:Getting hold of invidual evaulation can generate better model as author 
mentioned 35-40 % of wines  quality ratings is hard to predict by any models .
more observations of high quality wines can verified the trends .

> ** **:these are some complications, still trends can be drawn and three 
influential features were identified.it is very hard to make any determination
for exact factor behind good wines or bad wines with 100% accuracy .
There were 1600 readings in this red wines dataset.

> ** **: Multivariate charts, i felt challenged to interpret the charts with 
color and 2 axis. But once , i understand them well , it was easy to interpret.


> **References**:https://www.r-bloggers.com/using-r-correlation-heatmap-with-
ggplot2/
http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-
analyze-format-and-visualize-a-correlation-matrix-using-r-software
https://archive.ics.uci.edu/ml/datasets/wine+quality
https://onlinecourses.science.psu.edu/stat857/node/225/
https://sriramjaju.github.io/2017-09-10-EDA-RedWine/


